Persistence in functional topology: local-connectivity, q-tameness and a correction to a theorem of Morse

### 2021-04-09 introduction.tex
Add new abstract

The interplay between the critical set of a function and the topology of its domain is a cornerstone of modern mathematics.
Now a days, when thinking about the pioneering work of Morse in this interplay, a smooth function on a closed differential manifold is possibly our first association but, as we describe next, it should not be our last.
Morse Theory in the \textit{finite} context was masterfully presented in Milnor's famous book on the subject \cite{MR0163331}; where he also move beyond it to give a new proof of Bott's periodicity, applying Morse theory to the energy functional of paths in a Riemannian manifold.
Another example of the importance of Morse's insights in infinite settings is Floer's work on the Arnold conjecture.
Morse himself worked in a significantly less constraint context, publishing in the 1930s a series of papers \cite{Morse.1937, Morse.1938, Morse.1940, MR9102} in which he established the key results of Morse theory in the broad context of semi-continuous functionals on metric spaces.
He name this body of work \textit{functional topology}.
A motivating problem for him came from the study of minimal surfaces.
In 1931, Jesse Douglas solved Plateau’s Problem \cite{Douglas.1931} for which he received the fields medal in 1936.
The problem amounts to finding for a given admissible boundary curve $\Gamma$ an area-minimizing embedded surface $\Sigma$ with $\partial \Sigma = \Gamma$.
In \cite{Morse.1939, Morse.1941}, Morse and Tompkins used functional topology to give conditions for the existence of unstable minimal surfaces, i.e., critical points of the area functional that do not locally minimize area.

>>>>

The interplay between the critical set of a function and the topology of its domain is a cornerstone of modern mathematics.
Now a days, when thinking about the pioneering work of Marston Morse in this domain, a smooth function on a closed differential manifold is possibly our first association but, as we describe next, it should not be our last.
In the \textit{finite} context, Morse theory was masterfully presented in Milnor's famous book on the subject \cite{MR0163331}; where he also move beyond this setting to give a new proof of Bott's periodicity, applying Morse theory to the energy functional of paths in a Riemannian manifold.
Another example of the importance of Morse's insights in infinite contexts is Floer's work on the Arnold conjecture.
Morse himself worked in a significantly less constraint setup, publishing in the 1930s a series of papers \cite{Morse.1937, Morse.1938, Morse.1940, MR9102} in which he established the key results of Morse theory in the broad context of semi-continuous functionals on metric spaces.
He name this body of work \textit{functional topology}.
A motivating problem for him came from the study of minimal surfaces.
In 1931, Jesse Douglas solved Plateau’s Problem \cite{Douglas.1931} for which he received the fields medal in 1936.
The problem amounts to finding for a given admissible boundary curve $\Gamma$ an area-minimizing embedded surface $\Sigma$ with $\partial \Sigma = \Gamma$.
In \cite{Morse.1939, Morse.1941}, Morse and Tompkins used functional topology to give conditions for the existence of unstable minimal surfaces, i.e., critical points of the area functional that do not locally minimize area.

### 2021-04-09 main.tex
Changes before meeting
Split preliminaries
Add section: historical remarks
Consolidate sections connectivity, counterexample and theorem
Improve introduction
Updates from Overleaf
Add historical stuff from Max thesis
Add new abstract

\immediate\write18{bibtex \jobname}
\documentclass{amsart}
\usepackage{amsmath, amsthm, amssymb}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=blue]{hyperref}
\usepackage[capitalize,noabbrev]{cleveref}
\usepackage{tikz-cd}
%\usepackage{newtxtext,newtxmath}
\usepackage[numbers,sort,compress]{natbib}
\usepackage{graphicx}

@@

% commands
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\HS}{\mathrm{HS}}
\newcommand{\HT}{\mathrm{HT}}
\newcommand{\HLC}{\mathrm{HLC}}
\renewcommand{\H}{\mathbb{H}^d}
\newcommand{\dprime}{{\prime\prime}}
\DeclareMathOperator{\interior}{int}
\DeclareMathOperator{\im}{im}
\newcommand\CH{\check{H}}

@@

\title{Persistence in functional topology: q-tameness, local-connectivity and a correction to a theorem of Morse}
\author{}
\date{\today}

\begin{abstract}
With applications to the study of minimal surfaces in mind, Marston Morse wrote a series of papers developing versions of his famous inequalities for a broad class of real-valued functions on metric spaces. His work is remarkable in so far as it likely represents the first use of something that can be called persistent homology in the mathematical literature. From the modern point of view, it is particularly interesting that he uses Vietoris homology in order to enforce semi-continuity of persistent homology and that he states two topological conditions that supposedly ensure q-tameness. However, he is partly incorrect: We present a filtration that satisfies one of his conditions but has non-q-tame persistent homology. We also conjecture that his other condition, for which he provides no proof, is correct and discuss possible proof strategies.
\end{abstract}

\maketitle

\input{introduction}
\input{preliminaries}
\input{connectivity}
\input{counterexample}
\input{theorem}
\input{outlook}
%\input{other}

>>>>

\immediate\write18{bibtex \jobname}
\documentclass{amsart}
\usepackage{amsmath, amsthm, amssymb}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=blue]{hyperref}
\usepackage[capitalize,noabbrev]{cleveref}
\usepackage{tikz-cd}
%\usepackage{newtxtext,newtxmath}
\usepackage[numbers,sort,compress]{natbib}
\usepackage{graphicx}
\usepackage{csquotes}

@@

% commands
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\HS}{\mathrm{HS}}
\newcommand{\HT}{\mathrm{HT}}
\newcommand{\HLC}{\mathrm{HLC}}
\newcommand{\HE}{\mathbb{H}^d}
\renewcommand{\H}{\widetilde{H}}
\DeclareMathOperator{\interior}{int}
\DeclareMathOperator{\im}{im}
\newcommand\CH{\check{H}}
\newcommand{\Mod}[1]{#1-\mathbf{Mod}}

@@

\title[Persistence in functional topology]{Persistence in functional topology: local-connectivity, q-tameness and a correction to a theorem of Morse}
\author{}
\date{\today}

\begin{abstract}
	During the 1930's Morse developed a general theory, which he called functional topology, relating the critical set of a semi-continuous functional and the topology of its sublevel sets, and, in join work with Tompkins, used this theory to study minimal surfaces.
	Many of the key insights of functional topology have been rediscovered in the study of persistence homology.
	In particular, conditions enforcing the existence of an interval decomposition are as central to functional topology as they are to persistence theory.
	With modern applications to geometry in mind, in this article we present a comprehensive and up-to-date study of the most delicate of these conditions: local-connectivity. Providing, on passing, a correction to the Critical Surface Theorem of Morse and Tompkins.
\end{abstract}

\maketitle
\tableofcontents
\input{introduction}
\input{homology}
\input{persistence}
\input{connectivity}
%\input{outlook}
\input{historical}
%\input{other}
